\begin{abstract}

Building a universal multilingual automatic speech recognition (ASR) model that performs equitably across languages has long been a challenge due to its inherent difficulties.
To address this task we introduce a \textbf{L}anguage-\textbf{A}gnostic \textbf{M}ultilingual \textbf{A}SR pipeline through orthography \textbf{U}nification and language-specific \textbf{T}ransliteration (\shortname\footnote{Codes: https://github.com/sanghyang00/lama-ut}).
\shortname~operates without any language-specific modules while matching the performance of state-of-the-art models trained on a minimal amount of data.
Our pipeline consists of two key steps.
First, we utilize a universal transcription generator to unify orthographic features into Romanized form and capture common phonetic characteristics across diverse languages. 
Second, we utilize a universal converter to transform these universal transcriptions into language-specific ones.
In experiments, we demonstrate the effectiveness of our proposed method leveraging universal transcriptions for massively multilingual ASR.
Our pipeline achieves a relative error reduction rate of 45\% when compared to Whisper and performs comparably to MMS, despite being trained on only 0.1\% of Whisper's training data.
Furthermore, our pipeline does not rely on any language-specific modules.
However, it performs on par with zero-shot ASR approaches which utilize additional language-specific lexicons and language models.
We expect this framework to serve as a cornerstone for flexible multilingual ASR systems that are generalizable even to unseen languages.

\end{abstract}